Ultra Diffuse Galaxies associated with Milky-Way Analogs

### 2022-08-25 main.tex
Updates from Overleaf

\subsection{Abundance of UDGs and UPGs}\label{sec:n_udg}

\begin{figure*}
	\vbox{ 
		\centering
		\includegraphics[width=1\linewidth]{N_UDG_host_mass.pdf}
	}
    \caption{The abundance of UDGs as a function of host halo mass. \textit{Left}: We compile the UDG abundance measurements from literature covering a wide range of host mass. The UDG abundance of our work $N_{\rm UDG} = 0.76\pm 0.04$ is shown as red square, the power-law relation from \citet{vdBurg2017} is shown in gray, and our fitting result is shown in pink. Our UDG abundance is consistent with ELVES but significantly higher than that in SAGA. After including more measurements at the lower halo mass end, the power-law is shallower than the relation reported in \citet{vdBurg2017}. \textit{Right}: We split the UDG sample into bins based on the host halo mass (triangles) and $g-i$ color (pentagons). The UDG abundance is higher for more massive host and redder host. }
    \label{fig:n_udg}
\end{figure*}

@@

Many mechanisms are also proposed to explain the large size of UDGs in simulation. For isolated UDGs, burst stellar feedback can cause the stellar component of dwarf galaxy to expand \citep[e.g.,][]{DiCintio2017,Chan2018,Martin2019,Jiang2019}. Analytical models also support a scenario where UDGs originates from a population of dwarf galaxies with higher spin than normal dwarfs \citep{Amorisco2016,Rong2017,Liao2019}. For UDGs in groups and clusters, environmental effects are more pronounced and tidal interactions are believed to puff up normal dwarfs to become UDGs. \citet{Jiang2019} argue that normal dwarfs can be tidally heated near the orbital pericenter,

\citet{Jiang2019} find that the average color is redder when it's closer to the host, and no size gradient as a function of distance, also attributing that ram-pressure stripping quenches the UDGs in groups, rather than tidal stripping.

\citet{Tremmel2020} study UDG formation in cluster environment and find UDGs with higher mass are puffed up suddenly due to tidal heating at pericenter, UDGs with lower mass are gradually puffed up due to adiabatic expansion as a response to the mass loss (DM + gas + star) from tidal stripipng and ram pressures stripping.

@@

\section{Summary}

>>>>

\subsection{Abundance of UDGs and UPGs}\label{sec:n_udg}
\jiaxuan{Discuss NGVS}
\begin{figure*}
	\vbox{ 
		\centering
		\includegraphics[width=1\linewidth]{N_UDG_host_mass.pdf}
	}
    \caption{The abundance of UDGs as a function of host halo mass. \textit{Left}: We compile the UDG abundance measurements from literature covering a wide range of host mass. The UDG abundance of our work $N_{\rm UDG} = 0.76\pm 0.04$ is shown as red square, the power-law relation from \citet{vdBurg2017} is shown in gray, and our fitting result is shown in pink. Our UDG abundance is consistent with ELVES but significantly higher than that in SAGA. After including more measurements at the lower halo mass end, the power-law is shallower than the relation reported in \citet{vdBurg2017}. \textit{Right}: We split the UDG sample into bins based on the host halo mass (triangles) and $g-i$ color (pentagons). The UDG abundance is higher for more massive host and redder host. }
    \label{fig:n_udg}
\end{figure*}

@@

% How are UDGs puffed up?
Many mechanisms are also proposed to explain the large size of UDGs in simulation. For isolated UDGs, burst stellar feedback can cause the stellar component of dwarf galaxy to expand \citep[e.g.,][]{DiCintio2017,Chan2018,Martin2019,Jiang2019}. Analytical models also support a scenario where UDGs originates from a population of dwarf galaxies with higher spin than normal dwarfs \citep{Dalcanton1997,Amorisco2016,Rong2017,Liao2019}. \citet{Wright2021} show that field UDGs can also be formed from major mergers that cause star formation to
migrate outward. For UDGs in groups and clusters, environmental effects are more pronounced and tidal interactions are believed to puff up normal dwarfs to become UDGs. \citet{Jiang2019} use the NIHAO simulations and find that the UDGs in groups are either accreted as UDGs or accreted as normal dwarfs but get tidally heated near the orbital pericenter. \citet{Tremmel2020} study UDG formation in cluster environment and find UDGs with higher mass ($M_\star > 10^{8}\ M_\odot$) are puffed up suddenly due to tidal heating at pericenter, but UDGs with lower mass ($M_\star \sim 10^{7.5}\ M_\odot$) are gradually puffed up due to adiabatic expansion as a response to the mass loss from tidal stripping and ram pressures stripping.

For UDGs in groups, since they probably experienced strong tidal interactions, it is not obviously clear whether they are quenched by tidal stripping or ram pressure stripping. However, in the tidal stripping scenario, dark matter needs to be removed before a significant fraction of gas are deprived by tidal force, indicating a longer quenching timescale. \citet{Jiang2019} argue that if tidal stripping is dominant in quenching, the satellites closer to the host will have smaller size and stellar mass since stars are also stripped. They find no such trend in simulation, but find that the average color is redder when the UDGs closer to the host, supporting that ram pressure stripping is dominant. In Section \ref{sec:size_distr}, we split the sample into two radial distance bins, and we do not find significant change in size distribution as UDGs get closer to the host. We also do not find any significant dependence of stellar mass distribution on radial distance for both UDGs and UPGs. Together with the fact that the quenched fraction decreases as UDGs (UPGs) get closer to their hosts, our findings agree with \citet{Jiang2019} that ram pressure stripping is the major quenching channel for UDGs (UPGs) in MW analogs.

One surprising finding in this work is that although UPGs have extreme size given their stellar mass, they have similar quiescent fraction as normal satellites in MW analogs. In ELVES, Greene et al. (in prep) also find no trend of quenched fraction on the distance to average mass-size relation. This null trend hints that quenching and size growth must happen separately.

1. UPGs are accreted into MW hosts from fields as normal dwarfs. If they subsequently get quenched and puffed up together due to tidal forces, we will expect a higher quenched fraction for UPGs since they must undergo more violent tidal interactions. If UPGs are quenched due to ram pressure stripping and puffed up due to tidal heating (one time event at pericenter), they will also exhibit higher quenched fraction because ram pressure is also stronger at pericenter. If they are puffed up due to a longer term adiabatic expansion \citep{Tremmel2020} due to mass loss, then the expansion should also be proportional to the amount of gas loss, which translates to quenched fraction.

2. If UPGs are already puffed up before accretion (bursty SF, merger, high spin halo), but remain star-forming, then they are quenched together as other normal satellites.

Bursty SF can puff the size up, but can also quench the galaxy due to SN feedback.

Merger: early merger can redistribute the SF to the outer regions in UDGs, but doesn't change the total SFR too much. This might be a good thing.

High spin: blue field UDGs have higher spin.

Preprocessing in a lower-mass group can also puff up the UPGs. \citet{Samuel2022} show that preprocessing can boost the quenched fraction by 20\% but only at $M_\star \approx 10^8\ M_\odot$. For lower mass satellites, preprocessing merely change the quenched fraction. If preprocessing only increase the size of normal dwarfs to be UPGs but not quench them, it might be a plausible way to produce the observed trend in Fig. \ref{fig:qfrac}.

splashback: they must be quenched in their prior infall.

However, the efficiency of ram pressure stripping also depends on the stellar mass density which provides the restoring force. The diffuse stellar component makes UPGs more vulnerable to ram pressure stripping. \jiaxuan{But UDGs have higher HI density?}

\citet{Kado-Fong2021} study the intrinsic shape of UDGs in observation and find the shape agree with merger scenario for field UDGs \citep{Wright2021} and tidal heating/adiabatic expansion for group/cluster UDGs \citep{Jiang2019,Tremmel2020}.

@@

\section{Summary}
\begin{itemize}
    \item Future works: nucleation, intrinsic shape, LSBG cross-correlation, unsupervised machine learning, color gradient, non-par shapes.
\end{itemize}
